---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OmopSparkConnector

<!-- badges: start -->
[![R-CMD-check](https://github.com/oxford-pharmacoepi/OmopSparkConnector/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/oxford-pharmacoepi/OmopSparkConnector/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

OmopSparkConnector provides a Spark specific implementation of an OMOP CDM reference as defined by the omopgenerics R package. 

## Installation

You can install the development version of OmopSparkConnector from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("oxford-pharmacoepi/OmopSparkConnector")
```

## Creating a cdm reference using Sparklyr

Let's first load the R libraries.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(sparklyr)
library(OmopSparkConnector)
```

To work with OmopSparkConnector, we will first need to create a connection to our data using the sparklyr. In the example below, we have a schema called "omop" that contains all the OMOP CDM tables and then we have another schema where we can write results during the course of a study. We also set a write prefix so that all the tables we write start with this (which makes it easy to clean up afterwards and avoid any name conflicts with other users).

```{r, eval=FALSE}
con <- sparklyr::spark_connect(.....)
cdm <- cdmFromSpark(con,
  cdmSchema = "omop",
  writeSchema = "results",
  writePrefix = "study_1_"
)
```

For this introduction we'll use a mock cdm where we have a small synthetic dataset in a local spark database.
```{r}
cdm <- mockSparkCdm(path = file.path(tempdir(), "temp_spark"))
```

## Cross platform support

With our cdm reference created, we now a single object in R that represents our OMOP CDM data. 
```{r}
cdm
```

This object contains references to each of our tables

```{r}
cdm$person |>
  dplyr::glimpse()

cdm$observation_period |>
  dplyr::glimpse()
```

With this we can use familiar dplyr code . For example, we can quickly get a count of our person table.
```{r}
cdm$person |> 
  tally()
```

Behind the scenes, the dbplyr R package is translating this to SQL.
```{r}
cdm$person |> 
  tally() |> 
  show_query()
```

We can also make use of various existing packages that work with a cdm reference using this approach. For example, we can extract a summary of missing data in our condition occurrence table using the OmopSketch package.
```{r}
library(OmopSketch)
library(flextable)

missing_condition_data <- OmopSketch::summariseMissingData(cdm, "condition_occurrence")
tableMissingData(missing_condition_data, type = "flextable")
```

## Native spark support

As well as making use of packages that provide cross-platform functionality with the cdm reference such as OmopSketch, because OmopSparkConnector is built on top of the sparklyr package we can also make use of native spark queries. For example we can compute summary statistics on one of our cdm tables using spark functions.

```{r}
cdm$person |>
  sdf_describe(cols = c(
    "gender_concept_id",
    "year_of_birth",
    "month_of_birth",
    "day_of_birth"
  ))
```

With this we are hopefully achieving the best of both worlds. On the one hand we can participate in network studies where code has been written in such a way to work across database platforms. And then on the other we are able to go beyond this approach, writing bespoke code that makes use of Spark-specific functionality.

# Disconnecting from your spark connection 

We can disconnect from our spark connection like so
```{r}
cdmDisconnect(cdm)
```
